Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-02-23 paper.tex
Start of FastR

\begin{document}
\title{Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data}
\author{Valerie Hayot-Sasson}
\maketitle
\begin{abstract}
  Text of abstract.  
\end{abstract}
\tableofcontents
\chapter{Introduction}
intro here
\chapter{BigData and Open science in neuroimaging}
    \note{Likely will not be split in two separate sections}
    \section{large images}
        \begin{itemize}  
            \item BigBrain
            \item Micro CT 
            \item EM 
        \end{itemize}
    \section{large datasets}
    	\note{off the top of my head, but need to double check}
        \begin{itemize}
            \item HCP
            \item ADNI
            \item OpenFMRI
            \item BIDS/BIDS apps
        \end{itemize}
    An effective engine to process such data requires the following characteristics
        \begin{itemize}
            \item in-memory computing
            \item data locality
            \item lazy evaluation
        \end{itemize}
        \section{open-science /data sharing platforms for neuroimaging}
        		is open science just data sharing, computing. 
        		\subsection{Introduction}
		\subsection{CBRAIN}
		\subsection{OpenNEURO}
		\subsection{NeuroData}
\chapter{Workflow engines for neuroimaging data}
	\section{Definition of workflows/pipelines}
	\section{Nipype}
		\subsection{summary}
		Nipype\cite{nipype} is a popular neuroimaging pipelinening framework written in Python.
		It provides users with uniform access to the rich ecosystem of neuroimaging software
		libraries (e.g. SPM, FSL, Freesurfer) through its \textit{Interfaces}. \weird{Should users want to
		use their own custom tool, it is also possible to create an interface for that}. To execute 
		Python code in Nipype, the \textit{Function} interface can be used \tristan{for arbitrary code?}. The framework is 
		also easy to use, and thus does not contribute to the steep learning curve new
		researchers must face.

		A pipeline in Nipype, consists of a data analysis
                \textit{Workflow} of connected
                \textit{interfaces}. The workflow is a directed
                acyclic graph that connects the outputs of one
                interface to be the inputs of another. A workflow may
                also consist of connected workflows.

@@

        \note{pipeline description}
        A pipeline in PSOM, is a series of connected jobs, which can be
represented as a (static?) directed acyclic graph. The pipeline itself is implemented 
using the Matlab/Octave data type. Each job within the pipeline
is assigned a unique name. There is only one mandatory field in the job
description (command). The remaining four optional fields include the list input
files (files\_in), the list output files (files\_out), the list of files to be
deleted after execution (files\_clean) and any Octave or Matlab datatype
containing the remaining parameters necessary for job execution (opt).
Jobs in PSOM run in a protected environment where they only have access to the
parameters that are passed.

@@

        \note{pipeline execution}

@@

        \note{debugging and provenance}

@@

        \note{pipeline description}
        Pipelines in matlabbatch consist of a series of connected data
processing steps known as "modules" \note{DAG?? probably but not explicitly
listed anywhere}. All SPM modules, helper modules and their respective
dependencies can be defined using SPM's GUI \note{can probably also be done
through command-line}.

@@

        \note{pipeline execution}
        Pipeline execution can be achieved either through the provided GUI or by
the command-line interface. The GUI, however, requires significant user
interaction, as the user must repeat the "Runs" command as many times as
necessary to reexecute the command on multiple data. The CLI, however, avoids
this with the help of its \textit{spm\_jobman} interface.

@@

        \note{debugging and provenance}
        All batch scripts can be saved as a Matlab scripts. These scripts not
only contain the code that was executed, but also all input parameters. Should
any script fail in the process of executing, an error will be printed to the
Matlab console. If the user is using the GUI, the error output may also be saved
to a Matlab file.

@@

        \note{Pipeline description}
        A Nextflow pipeline is represented by the flow of data between processes. These
processes can be written in any Linux-compatible scripting language. As in Nipype, 
these processes are executed in an isolated environment. The only means
of communication between processes is through asynchronous FIFO queues known as
\textit{channels}. Each process can contain any number of input and output
channels. Through the use of channels, input and output can be stored not only
on disk but also in-memory \note{Nipype can also return outputs that are not on
disk, so this is nothing special}.

@@

        \note{Pipeline execution}
        The \textit{executor} is responsible for the execution of a Nextflow
pipeline. By default, it assumes that the pipeline is to be run locally. The
local executor will use multithreading in order to exploit available cores
within the local environment. To configure the executor to run in a remote distributed environment, it is
necessary to modify the configuration file to specify the environment. By
following this approach, Nextflow users can seamlessly run their code on various
processing platforms with a simple modification to a configuration file. In
addition to local, Nextflow currently has executors for SGE, LSF, SLURM,
PBS/Torque, NQSII, HTCondor, DRMAA, Ignite, Kubernetes and AWS Batch.

        \note{Debugging and Provenance}
        When executing a pipeline, it is always possible to specify the
generation of a execution report. The report is written in HTML and consists of three sections:
Summary, Resources and Tasks. Summary provides details on the execution command,
execution status, overall execution time, in addtion to other metadata. The
Resources tab, on the other hand, contains several HighChart plots representing
CPU, memory, time and disk read/write rate of each pipeline process. The final
section, Tasks, contains many metrics including the status and command script of
each task.

@@

    \section{FastR}
        \subsection{summary}
        \note{Pipeline description}
        \note{Pipeline execution}
        \note{Debugging and Provenance}
    \section{OpenMOLE}
        \subsection{summary}
        \note{Pipeline description}
        \note{Pipeline execution}
        \note{Debugging and Provenance}
    \section{LONI}
    \section{Automatic Analysis}
    \section{Dipy, Make, Pydiper} 
    \section{Use of containers in Workflow engines} 
        \subsection{use of containers in nipype, psom, spm}

@@

\chapter{Scientific workflows in Spark}
    \section{Introduction} 
    \section{ariel's paper and thunders astronomy - }
    \section{Spark on hpc's or } 
        \note{large dataset for valduriez} Prior to
examining how Big Data frameworks could be improved on HPC clusters, it is
important to have a baseline of how it performs natively on an HPC cluster.
Particularly, how it behave when processing scientific workflows. In a recent
study done by \cite{valduriez}, they examined the scalability of Spark on an HPC
cluster using a black-box scientific workflow. Five tests were performed in
their study: 1) Measuring execution time when task and resources available
increase proportionally, 2) Measuring the execution time of very short and short
tasks, 3) Measuring the execution time of short (5s) to long tasks (120s), 4)
Varying the number of tasks with fixed task durations, 5) Mixed task durations
and varied number of tasks. Their results show Spark behaves as expected when
tasks and resources are increased exponentially, and that Spark scales better
than expected with tasks of longer durations. This is due to the fact that the
scheduler is overloaded in the case of many short tasks in addition to their
being a significant amount of I/O occurring between each task. Although I/O is
problematic in both data center and HPC infrastructures, HPC is optimized for
data transfer over the network rather than disk I/O, and therefore, such
frequent I/O negatively impacts the system. However, many scientific workflows
are composed of longer tasks, and therefore, Spark scales very well with
scientific workflows in an HPC environment.

>>>>

\begin{document}
\title{Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data}
\author{Valerie Hayot-Sasson}
\maketitle
\begin{abstract}
  Text of abstract.  
\end{abstract}
\tableofcontents
\chapter{Introduction}
intro here
\chapter{BigData and Open science in neuroimaging}
    \note{Likely will not be split in two separate sections}
    \section{large images}
        \begin{itemize}  
            \item BigBrain
            \item Micro CT 
            \item EM 
        \end{itemize}
    \section{large datasets}
    	\note{off the top of my head, but need to double check}
        \begin{itemize}
            \item HCP
            \item ADNI
            \item OpenFMRI
            \item BIDS/BIDS apps
        \end{itemize}
    An effective engine to process such data requires the following characteristics
        \begin{itemize}
            \item in-memory computing
            \item data locality
            \item lazy evaluation
        \end{itemize}
        \section{open-science /data sharing platforms for neuroimaging}
        		is open science just data sharing, computing. 
        		\subsection{Introduction}
		\subsection{CBRAIN}
		\subsection{OpenNEURO}
		\subsection{NeuroData}
\chapter{Workflow engines for neuroimaging data}
	\section{Definition of workflows/pipelines}

    Broad definition of activities - between the tool and the engine.
	\section{Nipype}
		\subsection{summary}
		Nipype\cite{nipype} is a popular neuroimaging pipelinening framework written in Python.
		It provides users with uniform access to the rich ecosystem of neuroimaging software
		libraries (e.g. SPM, FSL, Freesurfer) through its \textit{Interfaces}. \weird{Should users want to
		use their own custom tool, it is also possible to create an interface for that}. To execute 
		Python code in Nipype, the \textit{Function} interface can be used \tristan{for arbitrary code?}. The framework is 
		also easy to use, and thus does not contribute to the steep learning curve new
		researchers must face.

		A pipeline in Nipype, consists of a data analysis. explicit
                \textit{Workflow} of connected
                \textit{interfaces}. The workflow is a directed
                acyclic graph that connects the outputs of one
                interface to be the inputs of another. A workflow may
                also consist of connected workflows.

@@

        \subsection{pipeline description}
        A pipeline in PSOM, is a series of connected jobs, which can be
represented as a (static?) directed acyclic graph. The pipeline itself is implemented 
using the Matlab/Octave data type. Each job within the pipeline
is assigned a unique name. There is only one mandatory field in the job
description (command). The remaining four optional fields include the list input
files (files\_in), the list output files (files\_out), the list of files to be
deleted after execution (files\_clean) and any Octave or Matlab datatype
containing the remaining parameters necessary for job execution (opt).
Jobs in PSOM run in a protected environment where they only have access to the
parameters that are passed.

@@

        \subsection{pipeline execution}

@@

        \subsection{debugging and provenance}

@@

        \subsection{pipeline description}
        Pipelines in matlabbatch consist of a series of connected data
processing steps known as "modules" \note{DAG?? probably but not explicitly
listed anywhere}. All SPM modules, helper modules and their respective
dependencies can be defined using SPM's GUI \note{can probably also be done
through command-line}.

@@

        \subsection{pipeline execution}
        Pipeline execution can be achieved either through the provided GUI or by
the command-line interface. The GUI, however, requires significant user
interaction, as the user must repeat the "Runs" command as many times as
necessary to reexecute the command on multiple data. The CLI, however, avoids
this with the help of its \textit{spm\_jobman} interface.

@@

        \subsection{debugging and provenance}
        All batch scripts can be saved as a Matlab scripts. These scripts not
only contain the code that was executed, but also all input parameters. Should
any script fail in the process of executing, an error will be printed to the
Matlab console. If the user is using the GUI, the error output may also be saved
to a Matlab file.

@@

        \subsection{Pipeline description}
        A Nextflow pipeline is represented by the flow of data between processes. These
processes can be written in any Linux-compatible scripting language. As in Nipype, 
these processes are executed in an isolated environment. The only means
of communication between processes is through asynchronous FIFO queues known as
\textit{channels}. Each process can contain any number of input and output
channels. Through the use of channels, input and output can be stored not only
on disk but also in-memory \note{Nipype can also return outputs that are not on
disk, so this is nothing special}.

@@

        \subsection{Pipeline execution}
        The \textit{executor} is responsible for the execution of a Nextflow
pipeline. By default, it assumes that the pipeline is to be run locally. The
local executor will use multithreading in order to exploit available cores
within the local environment. To configure the executor to run in a remote distributed environment, it is
necessary to modify the configuration file to specify the environment. By
following this approach, Nextflow users can seamlessly run their code on various
processing platforms with a simple modification to a configuration file. In
addition to local, Nextflow currently has executors for SGE, LSF, SLURM,
PBS/Torque, NQSII, HTCondor, DRMAA, Ignite, Kubernetes and AWS Batch.

        \subsection{Debugging and Provenance}
        When executing a pipeline, it is always possible to specify the
generation of a execution report. The report is written in HTML and consists of three sections:
Summary, Resources and Tasks. Summary provides details on the execution command,
execution status, overall execution time, in addtion to other metadata. The
Resources tab, on the other hand, contains several HighChart plots representing
CPU, memory, time and disk read/write rate of each pipeline process. The final
section, Tasks, contains many metrics including the status and command script of
each task.

@@

    \section{FastR}
        \subsection{summary}
        FastR is another Python-based neuroimaging pipeline framework that
provides added features such including access to different versions of the same
command-line tool \note{criticizes nipype for not having such a feature, but
nipype can access boutiques tools...so...} and compatibility with various
possible filesystems. In addition to its added features, FastR also guarantees
a simple-to-use framework for building pipelines written in any languages, has
built-in provenance support, provides the seamless ability to parallelize
pipelines with the use of plugins and avoids the numerical instability issue with
the simple integration of containerized pipelines. Similary to many other
pipeline framework, it also employs a DAG to describe data flow.

        \subsection{Pipeline description}
        A pipeline in FastR is a \textit{Network} of connected \textit{Nodes} in
which the directed edges represent the flow of data.  Each \textit{node}
contains a JSON or XML template known as a \textit{Tool}.

\textit{Tools}
describe the details of the underlying application that is to be executed at the
node. Tools consist of three components: general metadata, target and interface.
The general metadata provides details on the Tools such as name, version,
authors, etc, whereas the target specifies details on the execution environment.
The interface, on the other hand, provides details on the inputs and outputs of
the tool.

A JSON schema is required to validate the python object generated by the Tool. 
Fastr currently provides two schemas for object validation: the \textit{Tool} schema
and the \textit{FastrInterface} schema. It is, however, possible to add a custom
schema to FastR and define an \textit{Interface} for it.

The \textit{Interface} class is responsible for executing a call to the command line tool.
It dictates the tags that must be defined within the interface tag of the
\textit{Tool}. For instance, the generic interface class requires details on the
input and output's id, cardinality (number of input values) and DataType (a
python object denoting the details of the Input or Output's data type).
Verification is done by FastR to determine if the linked Input and Output data
types are compatible \note{data types can also be grouped, but did not really
understand that}.

In FastR, there are four types of \textit{Nodes}: normal Nodes, Sink Nodes,
Source Nodes and Constant Nodes. Data enters the network through Source or
Constant Nodes and exit through Sink Nodes. Constant nodes differ from Source
Nodes in that their input data are defined by the network \note{Not sure if
saying this correctly}. Normal nodes are responsible for processing the
data. Nodes are connected to each other through links that highlight the
dependencies and form the \textit{Network}. Once the \textit{Netwokr} is
defined, it can be executed.

        \subsection{Pipeline execution}
        \subsection{Debugging and Provenance}
    \section{OpenMOLE}
        \subsection{summary}
        \subsection{Pipeline description}
        \subsection{Pipeline execution}
        \subsection{Debugging and Provenance}
    \section{LONI}
    \section{Automatic Analysis}
    \section{Dipy, Make, Pydiper} 
    \section{Use of containers in Workflow engines} 
        \subsection{use of containers in nipype, psom, spm}

@@

\chapter{Scientific workflows in Spark}
    \section{Introduction}
    emphasis on data rather than on interface. transformation describe how
interfaces interact with the data. 
    \section{ariel's paper and thunders astronomy - }
    \section{Spark on hpc's or } 
        \note{large dataset for valduriez} Prior to
examining how Big Data frameworks could be improved on HPC clusters, it is
important to have a baseline of how it performs natively on an HPC cluster.
Particularly, how it behave when processing scientific workflows. In a recent
study done by \cite{valduriez}, they examined the scalability of Spark on an HPC
cluster using a black-box scientific workflow. Five tests were performed in
their study: 1) Measuring execution time when task and resources available
increase proportionally, 2) Measuring the execution time of very short and short
tasks, 3) Measuring the execution time of short (5s) to long tasks (120s), 4)
Varying the number of tasks with fixed task durations, 5) Mixed task durations
and varied number of tasks. Their results show Spark behaves as expected when
tasks and resources are increased exponentially, and that Spark scales better
than expected with tasks of longer durations. This is due to the fact that the
scheduler is overloaded in the case of many short tasks in addition to their
being a significant amount of I/O occurring between each task. Although I/O is
problematic in both data center and HPC infrastructures, HPC is optimized for
data transfer over the network rather than disk I/O, and therefore, such
frequent I/O negatively impacts the system. However, many scientific workflows
are composed of longer tasks, and therefore, Spark scales very well with
scientific workflows in an HPC environment.
